A Comparative Quality Metric for Untargeted Fuzzing with Logic State Coverage

### 2024-06-02 paper/main.tex
paper: discussion: quality metric for general fuzzing
intro, representativeness/normality

\title{A Comparative Quality Metric for Fuzzing with Logic State Coverage}
\author{
  {\rm Gwangmu Lee} \\ {EPFL}
}

@@

However, both metrics have been consistently refuted as unreliable. Edge
coverage has long been discredited as a \emph{weak} proxy metric as it
often leads to indecisive measurements; it is common to see the edge coverage
comparisons offering little difference (e.g., equally saturated edge coverage in
hypervisor virtual devices \cite{myung2022mundofuzz,sergej2021nyx}) or
influenced by initial inputs (e.g., different requirements for initial inputs in
grammar fuzzing \cite{chen2021polyglot}).
%, in which cases researchers should resort to another way for demonstrating effectiveness.  
Furthermore, some experiments yield counter-intuitive results where a code
coverage increment does not necessarily translate to more discovered bugs so
well \cite{google-ai}.

The number of discovered bugs, on the other hand, is a direct metric as it
counts the very objects (i.e., bugs) that fuzzing aims to reveal. However, it is
commonly known to be heavily sensitive to side factors such as bug
deduplication, initial inputs, and the implementation details of both the fuzzer
and the software under test \cite{klees2018evaluating}. It gets worse because
the number of total bugs is usually only a handful,
%regardless of the software under test, 
where unclear factors can introduce a measurement bias that significantly
exaggerates the result in certain software.

@@

As it is challenging to directly count interesting program behaviors when
fuzzing produces hundreds of them every second, we propose \emph{logic state
coverage} as a feasible proxy metric of interesting behaviors. A logic state is
a set of all satisfied branches during one execution, where logic state coverage
is the \emph{individual count} of all observed distinct logic states during a
fuzzing campaign. Logic states represent \emph{interesting} behaviors (i.e.,
relating to unknown abnormal behaviors) in a finer granularity, meaning that
observing more interesting behaviors is reliably reflected to larger coverage.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{images/local.pdf}
  \caption{Example control flow graph and program behaviors. The exceptional
  condition $e$ is satisfied when $\mathrm{mem}(q)$ is a freed memory. $l(X)$
  denotes the logic state of a program behavior $X$ (hit count omitted).}
  % Each region surrounded by dotted lines constitutes one logic state.}
  \label{f:local}
\end{figure}

@@

In one logic state, \emph{all} program behaviors are either normal or abnormal
altogether. 
%
This is because normal behaviors trigger \emph{exit branches} before
reaching program exit points, while abnormal behaviors do not. Since the logic
states of two behaviors differ by exit branches, they cannot have the same set
of satisfied branches, hence the same logic state.

\autoref{f:local} shows an example program that contains a use-after-free
exception $e$, which is satisfied when $\mathrm{mem}(q)$ belongs to freed
memories. Three example program behaviors and their logic states are shown on
the side. Since normal behaviors (Behaviors A and B) terminate at the program
exit, their logic states contain the exit branch $b_x$ that directly leads to
the program exit. On the other hand, since the abnormal behavior (Behavior C)
triggers the exception $e$ that crashes the execution before reaching the
program exit, its logic state does not contain any exit branch unlike normal
behaviors.

@@

To devise a quality metric that is also compatible to targeted fuzing, future
research may need to address two issues revolving around interestingness. First,
it should accurately describe the interestingness in an unpartial way (i.e., not
favoring a certain fuzzer or setting). To do so, it should assess which factor
affects interestingness and how much more than one another. Second, it should
incorporate such a description into the quality metric. Should it follow the
philisophy of logic state coverage, grouping uninteresting behaviors may be a
possible approach to make them less significant.

>>>>

\title{A Comparative Quality Metric for Untargeted Fuzzing with Logic State Coverage}
\author{
  {\rm Gwangmu Lee} \\ {iss300@gmail.com}
}

@@

However, both metrics have been consistently refuted as unreliable. Edge
coverage has long been discredited as a \emph{weak} proxy metric as it
often leads to indecisive measurements; it is common to see the edge coverage
comparisons offering little difference (e.g., equally saturated edge coverage 
%in hypervisor virtual devices 
\cite{myung2022mundofuzz,sergej2021nyx}) or
influenced by initial inputs (e.g., different requirements for initial inputs 
%in grammar fuzzing 
\cite{chen2021polyglot}).
%, in which cases researchers should resort to another way for demonstrating effectiveness.  
Furthermore, some experiments yield counter-intuitive results where a code
coverage increment does not translate to more discovered bugs so well
\cite{google-ai}.

The number of discovered bugs, on the other hand, is a much more direct metric
because it counts the very objects (i.e., bugs) that fuzzing aims to reveal.
However, it is highly sensitive to side factors such as bug deduplicatation
method, initial inputs, or the implementation details of both the fuzzer and the
software under test \cite{klees2018evaluating}. Even worse, the number of total
bugs is usually only a handful,
%regardless of the software under test, 
making unclear factors easily introduce a measurement bias that significantly
exaggerates the result in certain software.

@@

As it is challenging to directly count interesting program behaviors when
fuzzing produces hundreds of them every second, we propose \emph{logic state
coverage} as a feasible proxy metric of interesting behaviors. A logic state is
a set of all satisfied branches during one execution, encoding the
exhibited program behavior without branch repetition information.
As a consequence, less repetitive (i.e., more \emph{interesting}) program
behaviors are reflected as more \emph{distinct} logic states. Logic state
coverage counts all distinct observed logic states during a fuzzing campaign,
which means that a larger coverage suggests more observed interesting behaviors.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{images/local.pdf}
  \caption{Example control flow graph and program behaviors. The exceptional
  condition $e$ is satisfied when the memory pointed by $q$ is a freed
  memory. $l(X)$ denotes the logic state of a program behavior $X$.}
  % Each region surrounded by dotted lines constitutes one logic state.}
  \label{f:local}
\end{figure}

@@

As long as a program does not exhibit an exception at the program exit basic
block, all program behaviors in one logic state have the same normality,
namely either normal or abnormal altogether.
%
This is because normal behaviors always trigger the \emph{exit branches} that
lead to the exit blocks, while abnormal behaviors do not. Since the logic
states of two behaviors are always distinguished by exit branches, they cannot
have the same set of satisfied branches (i.e., the same logic state).

\autoref{f:local} shows an example program that contains a use-after-free
exception $e$, which is satisfied when the memory pointed by $q$ belongs to
freed memories. Three example program behaviors and their logic states are shown
on the side. Since normal behaviors (Behaviors A and B) terminate at the program
exit, their logic states contain the exit branch $b_x$ that directly leads to
the program exit. On the other hand, since the abnormal behavior (Behavior C)
triggers the exception $e$ that crashes the execution before reaching the
program exit, its logic state does not contain any exit branch unlike normal
behaviors.

@@

To devise a quality metric that also encompasses targeted fuzing, future
research may need to address two issues revolving around interestingness. First,
it should accurately describe the interestingness in an unpartial way (i.e., not
favoring a certain fuzzer or setting). To do so, it should assess which factor
affects interestingness and how much more than one another. Second, it should
incorporate such a description into the quality metric. Should it follow the
philisophy of logic state coverage, grouping uninteresting behaviors may be a
possible approach to make them less significant.
